Titeln på avhandlingens språk

### 2022-12-05 intro.tex
jaya

In that context, the contributions of this paper are as follows:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, data-driven model for human timings in \ac{WCA} applications.
    Using the data collected for \textcite{olguinmunoz:impact2021} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \ac{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguin2018scaling,olguin2019edgedroid}.
    \item\label{item:contrib:footprint} Using the model described above, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \ac{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz:impact2021}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \ac{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \ac{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Compared to the state-of-the-art, our optimization model results in up to \textasciitilde\SI{60}{\percent} fewer samples processed or an average of \SI{20}{\percent} less energy consumed per step, while maintaining comparable levels of responsiveness.
\end{enumerate}

>>>>

In that context, the contributions of this paper are as follows:
\begin{enumerate}
    \item\label{item:contrib:model} We introduce the first, to our knowledge, data-driven model for human timings in \ac{WCA} applications.
    Using the data collected for~\cite{olguinmunoz:impact2021} as a base, we build a stochastic model which takes as input past measurements of system responsiveness and produces realistic step execution times.
    We also introduce a novel way to generate dynamic traces of frames for \ac{WCA} applications which can be combined with the timing model for a full end-to-end emulation of a human.
    We name this new model \emph{\edgedroid}; a direct, more realistic evolution of our initial EdgeDroid approach~\cite{olguin2018scaling,olguin2019edgedroid}.
    \item\label{item:contrib:footprint} Using the model described above, we study the implications of realistic human behavior for the \emph{application lifetime footprint} of \ac{WCA}, understanding this term as the duration of a specific complete execution of the application task.
    In accordance with previous work~\cite{olguinmunoz:impact2021}, we find dependencies between system responsiveness and human step execution times that lead to substantially different application lifetimes when compared to a first-order baseline which does not take into account human behavior.
    \item\label{item:contrib:optimization} Finally, we study the potential for optimization in \ac{WCA} when considering human behavior using our model.
    We develop a generic model for the stochastic optimization of resource consumption versus responsiveness trade-offs in these applications, which we apply to two potential avenues for \ac{WCA} optimization; number of processed samples and energy consumption per step.
    First, we study the potential for reducing the number of samples captured per step.
    This is a valuable endeavor, as reducing the number of samples captured and subsequently processed directly translates in lower bandwidth demand on the wireless network and processor time demand on the cloudlet.
    Next, we explore the potential for direct optimization of energy consumption.
    The economic feasibility of WCA, and hence its likelihood of commercial adoption depends on it not being a resource hog, and this work furthers that effort.
    Compared to the state-of-the-art, our optimization model results in up to \textasciitilde\SI{60}{\percent} fewer samples processed or an average of \SI{20}{\percent} less energy consumed per step, while maintaining comparable levels of responsiveness.
\end{enumerate}

### 2022-12-05 model.tex
jaya

% As expressed in \cref{ssec:plos}, for \textcite{olguinmunoz:impact2021} we collected timing and task performance data for \num{40} participants performing a \num{169}-step task, guided by a \ac{WCA}, as well as personality trait scores for each participant.
% For this paper, we use this data to construct a probabilistic model for the generation of realistic execution times.
The processing of the data from \textcite{olguinmunoz:impact2021} for the generation of execution times can be summarized as a grouping according to discretized levels of neuroticism and a weighted rolling average of \ac{TTF}.
The resulting collections of execution times represent the distributions of these values for users with specific levels of neuroticism, interacting with systems at specific states of impairment and recent histories of impairment.
These distributions can then be sampled to produce new, realistic execution times.
In the following, we detail the step-by-step processing of this data and construction of the probabilistic model.

We employ a cleaned and re-parameterized copy of the timing data.
The \num{6760} data points are arranged in a table together with identifiers for the subjects, their neuroticism score, and a sequence number for each step.

\begin{table*}[]
    \centering
    \caption{%
        Default bins used for model parameter levels.
        Values have been rounded to two decimal places.
    }
    \label{tab:defaultbins}
    \begin{tabular}{@{}lccccccc@{}}
        \toprule
        \textbf{Parameter} & \textbf{Low} & & & \textbf{Medium} & & & \textbf{High}         \\ \midrule
        Neuroticism        & \( [0, 0.5) \) & & & & & & \( [0.5, 1.0] \)      \\
        Weighted TTF & \((-\infty, 0.82]\) & \((0.82, 1.53]\) & \((1.53, 2.08]\) & \((2.08, 2.67]\) & \((2.67, 3.45]\) & \((3.45, 4.13]\) & \((4.13, \infty]\)
    \end{tabular}
\end{table*}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \ac{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \ac{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bars indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:exectimeduration}
\end{figure}

>>>>

% As expressed in \cref{ssec:plos}, for \textcite{olguinmunoz:impact2021} we collected timing and task performance data for \num{40} participants performing a \num{169}-step task, guided by a \ac{WCA}, as well as personality trait scores for each participant.
% For this paper, we use this data to construct a probabilistic model for the generation of realistic execution times.
The processing of the data from~\cite{olguinmunoz:impact2021} for the generation of execution times can be summarized as a grouping according to discretized levels of neuroticism and a weighted rolling average of \ac{TTF}.
The resulting collections of execution times represent the distributions of these values for users with specific levels of neuroticism, interacting with systems at specific states of impairment and recent histories of impairment.
These distributions can then be sampled to produce new, realistic execution times.
In the following, we detail the step-by-step processing of this data and construction of the probabilistic model.

We employ a cleaned and re-parameterized copy of the timing data.
The \num{6760} data points are arranged in a table together with identifiers for the subjects, their normalized neuroticism score, and a sequence number for each step.

\begin{table*}[]
    \centering
    \caption{%
        Default bins used for model parameter levels.
        Values have been rounded to two decimal places.
    }
    \label{tab:defaultbins}
    \begin{tabular}{@{}lccccccc@{}}
        \toprule
        \textbf{Parameter} & \textbf{Low} & & & \textbf{Medium} & & & \textbf{High}         \\ \midrule
        Neuroticism        & \( [0, 0.5) \) & & & & & & \( [0.5, 1.0] \)      \\
        Weighted TTF & \([0.0], 0.82]\) & \((0.82, 1.53]\) & \((1.53, 2.08]\) & \((2.08, 2.67]\) & \((2.67, 3.45]\) & \((3.45, 4.13]\) & \((4.13, \infty]\)
    \end{tabular}
\end{table*}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \ac{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \ac{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bands indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:exectimeduration}
\end{figure}

### 2022-12-05 related_work.tex
jaya

There exist a number of approaches to the characterization and benchmarking of mobile \ac{XR}, and by extension \ac{WCA}, deployments.
\emph{OpenRTiST}~\cite{george2020openrtist} is a tool which uses a compute-heavy, yet latency-sensitive workload --- image style transfer~\cite{jing2019neural} --- to load and benchmark edge computing deployments.
It acts as an end-to-end realistic \ac{AR} workload, allowing thus researchers to measure the real end-to-end latencies of such systems.
\textcite{Chen2017Empirical} use a number of prototype \ac{WCA} implementation to study and characterize real-world latency bounds in these systems.
In \textcite{lecci2021open}, the authors collect and analyze large amounts of \ac{VR} network traffic, which they then use to construct a synthetic model for the generation of traces of such traffic.
\textcite{chetoui2022arbench} propose \emph{ARBench}, a toolkit for for the benchmarking of mobile hardware in the context of \ac{AR}.
The toolkit incorporates a series of workloads which stress different components of the mobile device and calculates a score for each workload.
\emph{Yahoo Cloud Serving Benchmark}~\cite{cooper2010benchmarking} and \emph{DCBench}~\cite{jia2013characterizing} are workload sets intended for benchmarking cloud services which have further been used in the context of benchmarking edge computing infrastructure for \ac{AR} suitability.
\emph{Edgebench}~\cite{das2018edgebench} and \emph{Defog}~\cite{mcchesney2019defog} benchmark workload performance on the edge versus on the cloud.

We contribute to this body of work by providing the first tool for the benchmarking of \ac{WCA} that dynamically considers the effects of human behavior.
In contrast to the above described works, which generate loads with relatively static profiles, our model is able to react to changes in system responsiveness like a human would.
We achieve this by building upon our own previous contributions to this field.
In \textcite{olguin2018scaling,olguin2019edgedroid} we introduced a coarse approximation to human behavior modeling in \ac{WCA} we called \emph{EdgeDroid}.
We used a trace-driven approach, where a pre-recorded and pre-processed ``ideal'' trace of steps for a specific \ac{WCA} step-based task is replayed to a Gabriel~\cite{Chen2018application} backend.
In order to adapt to potential mismatches between system responsiveness at trace-capture time and trace-replay time, we used a simple \ac{FSM} to adapt the trace at the latter by replaying or skipping certain segments.
In \textcite{olguinmunoz:impact2021} we studied the effects of reduced system responsiveness on human behavior in \ac{WCA} applications through human-subject studies.
We found that humans generally pace themselves according to the perceived system responsiveness.
When interacting with a highly responsive system, humans tended to speed up with each step; conversely, humans tended to slow down in highly unresponsive states.

The characterization and optimization, of mobile \ac{XR} has also long been a topic of research.
In \textcite{srinivasan2009performance}, the authors analyze a \ac{MAR} workload and identify and characterize the computational bottlenecks in the application.
They use this characterization to develop a series of code optimizations which allow them to achieve a threefold increase in performance.
The authors of \textcite{Wang2019Towards} develop a taxonomy for \ac{WCA} as well as strategies for workload reduction in these applications, including a novel adaptive sampling scheme with the goal of reducing the number of processed samples while maintaining application responsiveness.
In \textcite{huang2021proactive}, the authors present a framework for the proactive allocation of edge cloud resources while taking into account the inherent mobility of wearable and mobile \ac{AR}.
\textcite{al2017energy} propose a resource allocation strategy which leverages collaboration between \ac{AR} applications to minimize energy consumption.
As above, we have also previously directly contributed to this field.
In \textcite{Moothedath2021EnergyOptimal,Moothedath2022EnergyEfficient} we find the optimum periodic sampling interval which minimizes the energy tracking human progress of a specific subtask in \ac{WCA}.
This is extended into an aperiodic strategy in \textcite{Moothedath2022Aperiodic}.
We further develop this approach in this work by deriving a more practical approximate solution to the aperiodic strategy, which we then use to implement a novel framework for the generic optimization of responsiveness-resource consumption trade-offs.

>>>>

There exist a number of approaches to the characterization and benchmarking of mobile \ac{XR}, and by extension \ac{WCA}, deployments.
\emph{OpenRTiST}~\cite{george2020openrtist} is a tool which uses a compute-heavy, yet latency-sensitive workload --- image style transfer~\cite{jing2019neural} --- to load and benchmark edge computing deployments.
It acts as an end-to-end realistic \ac{AR} workload, allowing thus researchers to measure the real end-to-end latencies of such systems.
\textcite{Chen2017Empirical} use a number of prototype \ac{WCA} implementation to study and characterize real-world latency bounds in these systems.
In~\cite{lecci2021open}, the authors collect and analyze large amounts of \ac{VR} network traffic, which they then use to construct a synthetic model for the generation of traces of such traffic.
\textcite{chetoui2022arbench} propose \emph{ARBench}, a toolkit for for the benchmarking of mobile hardware in the context of \ac{AR}.
The toolkit incorporates a series of workloads which stress different components of the mobile device and calculates a score for each workload.
\emph{Yahoo Cloud Serving Benchmark}~\cite{cooper2010benchmarking} and \emph{DCBench}~\cite{jia2013characterizing} are workload sets intended for benchmarking cloud services which have further been used in the context of benchmarking edge computing infrastructure for \ac{AR} suitability.
\emph{Edgebench}~\cite{das2018edgebench} and \emph{Defog}~\cite{mcchesney2019defog} benchmark workload performance on the edge versus on the cloud.

We contribute to this body of work by providing the first tool for the benchmarking of \ac{WCA} that dynamically considers the effects of human behavior.
In contrast to the above described works, which generate loads with relatively static profiles, our model is able to react to changes in system responsiveness like a human would.
We achieve this by building upon our own previous contributions to this field.
In~\cite{olguin2018scaling,olguin2019edgedroid} we introduced a coarse approximation to human behavior modeling in \ac{WCA} we called \emph{EdgeDroid}.
We used a trace-driven approach, where a pre-recorded and pre-processed ``ideal'' trace of steps for a specific \ac{WCA} step-based task is replayed to a Gabriel~\cite{Chen2018application} backend.
In order to adapt to potential mismatches between system responsiveness at trace-capture time and trace-replay time, we used a simple \ac{FSM} to adapt the trace at the latter by replaying or skipping certain segments.
In~\cite{olguinmunoz:impact2021} we studied the effects of reduced system responsiveness on human behavior in \ac{WCA} applications through human-subject studies.
We found that humans generally pace themselves according to the perceived system responsiveness.
When interacting with a highly responsive system, humans tended to speed up with each step; conversely, humans tended to slow down in highly unresponsive states.

The characterization and optimization of mobile \ac{XR} has also long been a topic of research.
\textcite{srinivasan2009performance} analyze a \ac{MAR} workload and identify and characterize the computational bottlenecks in the application.
They use this characterization to develop a series of code optimizations which allow them to achieve a threefold increase in performance.
\textcite{Wang2019Towards} develop a taxonomy for \ac{WCA} as well as strategies for workload reduction in these applications, including a novel adaptive sampling scheme with the goal of reducing the number of processed samples while maintaining application responsiveness.
\textcite{huang2021proactive} present a framework for the proactive allocation of edge cloud resources while taking into account the inherent mobility of wearable and mobile \ac{AR}.
\textcite{al2017energy} propose a resource allocation strategy which leverages collaboration between \ac{AR} applications to minimize energy consumption.
As above, we have also previously directly contributed to this field.
In~\cite{Moothedath2021EnergyOptimal,Moothedath2022EnergyEfficient} we find the optimum periodic sampling interval which minimizes the energy tracking human progress of a specific subtask in \ac{WCA}.
This is extended into an aperiodic strategy in~\cite{Moothedath2022Aperiodic}.
We further develop this approach in this work by deriving a more practical approximate solution to the aperiodic strategy, which we then use to implement a novel framework for the generic optimization of responsiveness-resource consumption trade-offs.
